---
title: "Making Figures"
output: rmarkdown::html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)

starttime <- Sys.time()

library(webmorphR)
wm_opts(plot = "ggplot")
wm_opts(plot.maxwidth = 850*2)
```


WebmorphR has plotting functions to make it easier to create reproducible figures from your stimuli.

When working interactively in an R Markdown document, the figures plot in the Viewer window, but will be shown in the document when you knit.

```{r plot}
stimuli <- demo_stim()
plot(stimuli)
```


## Borders and Background

You can add a border to each image using pad. Values less than 1 will set border width as a percentage of the image width, while values greater than 1 will set it as pixels. You can also change the border colour using `fill`.


```{r pink}
stimuli %>%
  pad(0.05, fill = "hotpink") %>%
  plot()
```


You can set different border sizes for the top, right, bottom and left sides (think "TRouBLe"). This can be useful for creating space for labels outside the image.

```{r border}
stimuli %>%
  pad(40, 0, 0, 0, fill = "black") %>%
  label(color = "white") %>%
  plot()
```


You can set the border colour separately for each image.

```{r}
stimuli %>%
  pad(10, fill = c("#F9DEC9", "#BAEDDD")) %>%
  plot()
```

## Rows and columns

Set `nrow`` or `ncol` to control the number of rows and columns.

```{r}
comp <- demo_stim("composite") %>% resize(0.5)

plot(comp, nrow = 2)
```


To make a multipart figure, you usually don't want padding around the outside of an image, so set `external_pad = FALSE`.  Set `byrow = FALSE` to distribute images by column.

```{r}
ind <- comp[c(1:2,4:7,9:10)] %>%
  plot(ncol = 2, byrow = FALSE, padding = 60, external_pad = FALSE)

c(comp[3], ind, comp[8]) %>%
  resize(height = 500) %>% # make sure parts are the same height
  plot(nrow = 1)
```


## Delineations

Use the `draw_tem()` function to show the delineations. By default, points are translucent green circles and lines are translucent blue.

```{r delin}
stimuli %>%
  draw_tem() %>%
  plot()
```

You can change the default colours and translucency with the `line.color` `pt.color` and `line.alpha` and `pt.alpha` arguments. Remove the image and set the background to grey by setting `bg = "grey"`.

```{r delin-custom}
stimuli %>%
  resize(1000) %>%
  draw_tem(line.color = "grey", line.size = 2, line.alpha = 0.2, 
           pt.color = "red", pt.size = 4, bg = "grey") %>%
  crop_pad(50) %>%
  plot()
```


## Delete some points

```{r point-remove}
remove_points <- frl_features("imprecise")
stimuli %>%
  pt_delete(remove_points) %>%
  draw_tem() %>%
  plot()
```

## Labels

Label figure panels with the image name.

```{r}
stimuli %>%
  label() %>%
  plot()
```

Use the `setnames()` function to set new names or set labels to a vector of labels to use.

```{r}
demo_stim("composite") %>%
  setnames(pattern = "f_", replacement = "Female ") %>%
  setnames(pattern = "m_", replacement = "Male ") %>%
  label() %>%
  plot(nrow = 2)
```

Labels can be positioned with the `gravity` and `location` arguments from the {magick} package.

```{r custom-labels}
stimuli %>%
  label(
    c("Female", "Male"),
    gravity = "northwest",
    location = "+10+10",
    color = c("darkgreen", "darkmagenta"),
    size = 20, 
    style = c("normal", "italic")
  ) %>%
  plot()
```

You can also get more fine control of label position.

```{r label-position}
stimuli %>%
  label(TRUE,
        color = "dodgerblue3",
        size = 20,
        weight = 700,
        gravity = "southwest",
        location = "+10+50",
        degrees = 45
  ) %>%
  plot()
```




```{r, echo = FALSE}
elapsed <- (Sys.time() - starttime) %>%
  as.numeric(units="mins") %>%
  round(1)
```

This script took `r elapsed` minutes to render all the included images from scratch.
