---
title: "Image Similarity"
output: rmarkdown::html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)

starttime <- Sys.time()

library(webmorphR)
wm_opts(plot.maxwidth = 850)
library(ggplot2)
library(tidyr)
```

This article explores the use of {magick}'s image_compare function in the context of morphed images, since I couldn't find any clear guides to what it was doing. I've implemented this for webmorph stimuli with the `image_comp()` function.

While this metric does capture something about similarity between morphed images, it is unlikely to be of any use as a metric for similarity between non-morphed images (e.g., it won't be able to tell you if two face images are of the same person or not).

### Create images to compare

```{r, fig.width = 11, fig.height = 1.5}
stimuli <- demo_stim()
s <- continuum(stimuli[1], stimuli[2]) %>%
  mask("face") %>%
  subset_tem(features("face")) %>%
  crop_tem()

plot(s, nrow = 1)
```

### Calculate metrics

Using all the available metrics, calculate similarity of each image to the first image.

```{r}
mtypes <- magick::metric_types()
names(mtypes) <- mtypes
m <- lapply(mtypes, image_comp, 
            stim = s, 
            ref_stim = s[[1]], 
            fuzz = 0, 
            scale = TRUE) # results scaled from 0 to 1
```


```{r, echo = FALSE}
df <- as.data.frame(m)
df$morph_distance <- seq(0,100,10)

df <- pivot_longer(df, cols = all_of(mtypes),
                 names_to = "metric",
                 values_to = "value")

ggplot(df, aes(morph_distance, value, color = metric)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~metric, nrow = 3)
```

PSNR is `Inf` when the test and reference image are identical, so the highest non-infinite value scales as 1.0. 

### Check fuzz settings

```{r}
m2 <- lapply(mtypes, image_comp, 
             stim = s, 
             ref_stim = s[[1]], 
             fuzz = 50,
             scale = TRUE)
```

Fuzz only does something for AE.

```{r, echo = FALSE}
df2 <- as.data.frame(m2)
df2$morph_distance <- seq(0,100,10)

df2 <- pivot_longer(df2, cols = all_of(mtypes),
                 names_to = "metric",
                 values_to = "value")

df$fuzz = "0"
df2$fuzz = "50"

dplyr::bind_rows(df, df2) %>%
  ggplot(aes(morph_distance, value, color = fuzz, shape = fuzz)) +
  geom_point() +
  facet_wrap(~metric, nrow = 3) +
  scale_shape_manual(values = c(16, 4))
```


```{r}
fuzzes <- seq(0, 20, 1)
names(fuzzes) <- fuzzes
m3 <- lapply(fuzzes, image_comp, 
             stim = s, 
             ref_stim = s[[1]], 
             metric = "AE",
             scale = FALSE) # don't scale for this comparison
```

```{r, echo = FALSE}
df3 <- as.data.frame(m3)
names(df3) <- fuzzes
df3$morph_distance <- seq(0,100,10)

df3 <- pivot_longer(df3, cols = 1:length(fuzzes),
                 names_to = "fuzz",
                 values_to = "value") %>%
  readr::type_convert()

ggplot(df3, aes(morph_distance, value, color = fuzz, group = fuzz)) +
  geom_point(show.legend = T) +
  geom_line(show.legend = F) +
  scale_color_viridis_c()
```


### Conclusions

* Increases with morph distance:
    * very strong negative exponential decay at 0 fuzz; more linear with higher fuzz: AE
    * strong negative exponential decay: PAE
    * slight negative exponential decay: Fuzz, RMSE
    * linear: MAE, MEPP, MSE
    * no idea: PHASH
* Decreases with morph distance:
    * linear: NCC, Undefined
    * slight exponential decay: PSNR
    

