---
title: "Making Stimuli"
output: rmarkdown::html_document
resource_files:
  - vignettes/images/stimuli/*
opengraph:
  image: 
    src: "https://debruine.github.io/webmorphR/articles/images/stimuli/stimuli.png"
  twitter:
    card: summary
    creator: "@lisadebruine"
    site: "@webmorph_org"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.path = "images/stimuli/"
)
starttime <- Sys.time()
library(dplyr)
```


This vignette will show some recipes for common types of stimulus creation.

```{r, message=FALSE}
# load packages and set maximum plot width
library(webmorphR)
wm_opts(plot.maxwidth = 850)
```


## Averageness

### Load demo images

Let's start with a set of images from the demo image sets. You need to have the [webmorphR.stim](https://github.com/debruine/webmorphR.stim) package installed if you haven't done that already.

```{r, eval = FALSE}
remotes::install_github("debruine/webmorphR.stim")
```

### Add info

You can add data to an image set to use in subsetting the images. For demonstration, the "london" image set has a data table called `london_info`.

```{r}
head(webmorphR.stim::london_info)
```

Use `add_info()` to add the data. If the data table isn't in the same order as the stimuli, you can match the stimulus item names to a data column using `.by`.

```{r add-info}
stimuli <- demo_stim("london") |>
  add_info(webmorphR.stim::london_info)
```

### Subset

Select images from your stimuli to make a new image subset. The original images are large, and we don't need the resulting stimuli to be that large, so we'll reduce image size by 50% right at the start to reduce processing time.

```{r subset}
subset <- stimuli |>
  subset(face_eth == "black") |>
  subset(face_gender == "male") |>
  resize(.5)

plot(subset, nrow = 2)
```

### Align

Next, align the images using Procrustes normalisation to the position of the first image. They need to have templates fitted to do this. The argument `patch = TRUE` tries to match the background colour as closely as possible. 

The `horiz_eyes` function just makes sure that the first image has good alignment, since his head is slightly tilted.

```{r align, message=FALSE}
aligned <- subset |>
  horiz_eyes(patch = TRUE) |>
  align(procrustes = TRUE, patch = TRUE)

plot(aligned, nrow = 2)
```

**Note**: If you get an error message about rgl or dynlib, and are using a Mac, you may need to install [XQuartz](https://www.xquartz.org/).

### Crop

You may also want to crop the images to a 3x4 aspect ratio by setting width to 60% of the image and height to 80%. By default, `crop()` will center the cropping, but you can also manually set the x and y offsets. Values less than 2.0 are interpreted as percentages, while values greater than 2.0 are interpreted as pixels. 

```{r crop}
cropped <- crop(aligned, width = 0.6, height = 0.8, y_off = .05)

plot(cropped, nrow = 2)
```

### Average

Now we can make an average version of these faces. This uses the morphing functions available on the web app, so you need to have an internet connection. It usually takes 1-4 seconds per image to upload your images to the server for processing.

```{r average}
avg <- avg(cropped)

# show average with and without template
draw_tem(avg) |> c(avg) |> plot()
```

### Transform

You can use this average face to transform the individual faces in distinctiveness and averageness. Give the shape vector names to set the output names automatically.

```{r transform}
transf <- trans(trans_img = cropped,
                from_img = avg,
                to_img = cropped,
                shape = c(avg = -0.5, dist = 0.5))
```

Use `subset()` to plot by row.

```{r male-grid}
plot_rows(
  "More Average" = subset(transf, "avg"),
  "Original" = cropped,
  "More Distinctive" = subset(transf, "dist"),
  top_label = TRUE
)
```

### Save stimuli

Now you can set names and save your individual stimuli in a directory to use in studies. The code below will create a directory called "stimuli" in your working directory if there isn't one already. You can write to any format that `magick::image_write()` handles, such as "png", "jpeg", or "gif" (the default is "png").

```{r save-avg}
avg |> write_stim(dir = "stimuli", 
                   names = "m_avg", 
                   format = "jpg")
```

Use `setnames()` to search and replace patterns, add prefixes or suffixes, or replace the names with a new vector of names. The London images in the `stimsets` package have "_03" appended to designate that they are the neutral front version from the [larger set](https://figshare.com/articles/dataset/Face_Research_Lab_London_Set/5047666), which also includes smiling images and several different viewpoints.

```{r setnames}
cropped |>
  setnames(pattern = "_03", 
           replacement = "", 
           prefix = "orig_") |>
  write_stim(dir = "stimuli", format = "jpg")

transf |>
  setnames(pattern = "_03", replacement = "") |>
  write_stim(dir = "stimuli", format = "jpg")
```

### Repeat

Now you can pipe all of the commands together and apply them to a new set of images, such as all the black women in the set.

```{r}
# subset, resize, align, and crop
cropped <- stimuli |>
  subset(face_gender == "female") |>
  subset(face_eth == "black") |>
  resize(.5) |>
  horiz_eyes(patch = TRUE) |>
  align(procrustes = TRUE, patch = TRUE) |>
  crop(width = 0.6, height = 0.8, y_off = 0.05)

# average
avg <- avg(cropped)

# transform
transf <- trans(trans_img = cropped,
                from_img = avg,
                to_img = cropped,
                shape = c(avg = -0.5, dist = 0.5))

# save
avg |> write_stim("stimuli", "f_avg", "jpg")

cropped |>
  setnames(pattern = "_03", 
           replacement = "", 
           prefix = "orig_") |>
  write_stim("stimuli", format = "jpg")

transf |>
  setnames(pattern = "_03", replacement = "") |>
  write_stim("stimuli", format = "jpg")

```

```{r female-grid}
plot_rows(
  "More Average" = subset(transf, "avg"),
  "Original" = cropped,
  "More Distinctive" = subset(transf, "dist"),
  top_label = TRUE
)
```

## Sexual Dimorphism

### Read in images

To manipulate sexual dimorphism, you need male and female average faces like the ones created above. Load them from the saved files in the "stimuli" directory.

```{r}
avgs <- read_stim(path = "stimuli", pattern = "(f|m)_avg")
orig <- read_stim(path = "stimuli", pattern = "orig_")
```

### Transform

You can then transform them by -50% to feminise and +50% to masculinise. Remember to give your shape vector names to set the output names automatically.

```{r sexdim}
sexdim <- trans(trans_img = orig[1:6],
                from_img = avgs$f_avg,
                to_img = avgs$m_avg,
                shape = c(fem = -0.5, masc = 0.5))

plot(sexdim, nrow = 2)
```


## Animate

### Make a continuum

Make a continuum that morphs from the female to the male average in 5% steps. This makes 21 images, so will take a few seconds; be patient.

```{r}
continuum <- continuum(avgs$f_avg, avgs$m_avg, 0, 1, .05)

plot(continuum, nrow = 3)
```

### Animate

You can turn your images into an animated gif. Resize the images to the size you want first. Save the resulting image as a gif. If you have the `gifski` package installed, this will be used to create the animated gif and is faster than the `magick` algorithm (you don't have to load `gifski` for it to be used, just install it).

```{r animated-gif, out.width = "180px"}
#install.packages("gifski") # for faster animation

continuum |>
  resize(width = 180) |>
  animate(fps = 20, rev = TRUE)
```

When you run this code, even in an R Markdown document, the gif will show in the Viewer pane, but it will display in the document when you knit.


## Non-Face Stimuli

You can also process images without templates. For example, the following code takes a group of images and crops them to a standard size.

```{r, message = FALSE}
# load rainbow images
stimuli <- demo_stim("rainbow")

# get info on the images to put in order by type and colour
info <- system.file("rainbow/_info.csv", package = "webmorphR.stim") |>
  readr::read_csv() |>
  dplyr::mutate(colour = factor(colour, c("red", "orange", "yellow", "green", "blue", "purple"))) |>
  dplyr::arrange(type, colour)

# crop to smallest size
width <- width(stimuli, "min")
height <- height(stimuli, "min")
stim <- crop(stimuli, width, height)
stim <- stim[info$photo_name] # reorder by type and colour

plot(stim, ncol = 6)
```

## Blank stimuli

You can make blank images specifying the size and colour.

```{r blank-stim}
new_stimuli <- blank(n = 6, width = 100, height = 100, 
                     color = c("red", "orange", "yellow", "green", "blue", "purple"))

plot(new_stimuli, nrow = 1)
```

## Word Stimuli

Create word stimuli by starting with blank images and adding words to each stimulus with the `label()` function.

```{r stroop-words}
# make a vector of the words and colours they should print in
colours <- c(red = "red3", 
             orange = "darkorange", 
             green = "darkgreen", 
             blue = "dodgerblue3", 
             purple = "darkorchid")

# make vector of labels (each word in each colour)
n <- length(colours)
labels <- rep(names(colours), each = n)

# make 36 blank 800x200px images and add labels
stroop <- blank(n*n, 800, 200) |>
  label(labels, 
        size = 100, 
        color = colours, 
        weight = 700,
        gravity = "center")

plot(stroop, ncol = n, fill = "grey")
```




```{r, echo = FALSE}
elapsed <- (Sys.time() - starttime) |>
  as.numeric(units="mins") |>
  round(1)
```

This script took `r elapsed` minutes to render all the included images from scratch.


```{r, echo = FALSE, eval = FALSE}
# cleanup
unlink("stimuli", recursive = TRUE)
```


