---
title             : "Reproducible Methods for Face Research"
shorttitle        : "Face Methods"

author: 
  - name          : "Lisa DeBruine"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "62 Hillhead Street, Glasgow G12 8QB, Scotland"
    email         : "lisa.debruine@glasgow.ac.uk"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Funding acquisition
      - Methodology
      - Software
      - Validation
      - Visualization
      - Writing - Original Draft Preparation
  - name          : "Iris Holzleitner"
    affiliation   : "1,2"
    role:
      - Software
      - Validation
      - Visualization
      - Writing - Review & Editing
  - name          : "Benedict C. Jones"
    affiliation   : "3"
    role:
      - Conceptualization
      - Writing - Review & Editing
affiliation:
  - id            : "1"
    institution   : "Institute of Neuroscience & Psychology, University of Glasgow"
  - id            : "2"
    institution   : "University of the West of England, Bristol"
  - id            : "3"
    institution   : "University of Strathclyde, Glasgow"

authornote: |
  This research was funded by ERC grant #647910 (KINSHIP).

abstract: |
  Face stimuli are commonly created in non-reproducible ways. This paper will introduce the open-access online platform webmorph and its associated R package webmorphR. It will explain the technical processes of morphing and transforming through a case study of creating face stimuli from an open-access image set.
  
keywords          : "faces; morphing; transforming; reproducible; webmorph"
wordcount         : "X"

bibliography      : ["r-references.bib","biblio.bib"]
csl               : apa.csl
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : yes

documentclass     : "apa6"
classoption       : "man"
output            : 
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: false
    self_contained: false
    code_folding: hide
note              : "Draft"
---

```{r setup, include = FALSE}
# devtools::install_github("crsh/papaja")
library("papaja")
library("kableExtra")
library("dplyr")
# devtools::install_github("debruine/webmorphR")
library("webmorphR") 
r_refs("r-references.bib")

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache       = TRUE,
  cache.extra = knitr::rand_seed,
  out.width   = "100%",
  warning     = FALSE,
  message     = FALSE,
  echo        = FALSE
)

wm_opts(plot.maxwidth = 850)
```

<style>
  blockquote { 
    font-size: 13px; 
    border: 1px solid grey; 
    background-color: #EEE;
    border-radius: 1em;
  }
  
  address { display: none; }
</style>

## Introduction

Face stimuli are commonly used in research on visual and social perception. This almost always involves some level of stimulus preparation to rotate, resize, crop, and reposition faces on the image. In addition, many studies systematically manipulate face images by changing color and/or shape properties [e.g., @jones2019biological; reviewed in @Little_2011]. 

Gronenschild and colleagues [-@Gronenschild_2009] argue for the importance of standardizing face stimuli so that they are not "confounded by factors such as brightness and contrast, head size, hair cut and color, skin color, and the presence of glasses and earrings". They describe a three-step standardization process. First, they manually removed features such as glasses and earrings in Photoshop. Second, they geometrically standardized images by semi-automatically defining eye and mouth coordinates used to fit the images within an oval mask, Third, they optically standardized images by converting them to greyscale and remapping values between the minimum and 98% threshold onto the full range of values. While laudable in its aims, this procedure has not achieved widespread adoption, probably because the authors provided no code or tools. In personal communication, the main author said that this is because "the procedure is based on standard image processing algorithms described in many textbooks". However, I was unable to easily replicate the procedure and found several places where instructions had more than one possible interpretation or relied on the starting images having specific properties, such as symmetric lighting reflections in the eyes.

The goal of this paper is to argue for the imortance of reproducible stimulus processing methods in face research and to introduce an open-source R package that allows researchers to create face stimuli with a script that can then be shared so that others can create stimuli using identical methods.

**Scope of this type of research. How many papers, what range of questions?**

### Reproducibility!

Why are reproducible stimulus construction methods important?

I once gave up on a research project because I couldn't figure out how to manipulate spatial frequency in MatLab to make my stimuli look like those in a relevant paper. When I contacted the author, they didn't know how the stimuli were created because a postdoc had done it in Photoshop and didn't leave a detailed record of the method.

Reproducibility of stimuli is especially important for face stimuli because faces are sampled, so replications should sample new *faces* as well as new participants [@barrgeneralizing]. The difficulty of creating equivalent face stimuli is a major barrier to this, resulting in stimulus sets that are used across dozens or hundreds of papers. For example, the Chicago Face Database [@CFD_2015] has been cited in almost 800 papers. Ekman and Friesen's [-@ekman1976pictures] Pictures of Facial Affect has been cited more than 5500 times. This image set is currently [**selling** for $399](https://www.paulekman.com/product/pictures-of-facial-affect-pofa/) for "110 photographs of facial expressions that have been widely used in cross-cultural studies, and more recently, in neuropsychological research". Such extensive reuse of image sets means that any confounds present in the image set can cause highly "replicable" but false findings.

Additionally, image sets are often private and reused without clear attribution. The Face research Lab has only recently been trying to combat this by making image sets public and citable where possible [e.g., @FRL_London;@Canada2003;@Morrison_2018] and including clear explanations of reuse where not possible [e.g., @jones2018no,@holzleitner2019comparing].

### Common Techniques

It was basically impossible to systematically survey the literature about the methods used to create facial stimuli, in large part because of poor documentation. However, several common methods are discussed below.

#### Photoshop/Image editors

A search for "Photoshop face attractiveness" produced 6,450 responses in Google Scholar. Here are descriptions of the use of Photoshop from a few of the top hits.

> If necessary, scanned pictures were rotated slightly, using Adobe Photoshop software, clockwise to counterclockwise until both pupil centres were on the same y-coordinate. Each picture was slightly lightened a constant amount by Adobe Photoshop. [@Scheib_1999, p. 1914]

> These pictures were edited using Adobe Photoshop 6.0 to remove external features (hair, ears) and create a uniform grey background. [@sforza2010my, pp. 150] 

  
#### Commerical morphing 

Face averaging or "morphing" is a common technique for making images that are blends of two or more faces. We found 831 Google Scholar responses for "Fantamorph face", 158 Google Scholar responses for "WinMorph face" and fewer mentions of several other programs, such as MorphThing (no longer available) and xmorph.

Most of these programs do not use open formats for storing delineations, the x- and y-coordinates of the landmark points that define shape and the way these are connected with lines. Their algorithms also tend to be closed and there is no common language for describing the procedures used to create stimuli in one program in a way that is easily translatable to another program. Here are descriptions of the use of commercial morphing programs from a few of the top hits.

> The faces were carefully marked with 112 nodes in FantaMorph™, 4th version: 28 nodes (face outline), 16 (nose), 5 (each ear), 20 (lips), 11 (each eye), and 8 (each eyebrow). To create the prototypes, I used FantaMorph Face Mixer, which averages node locations across faces. Prototypes are available online, in the Personality Faceaurus [http://www.nickholtzman.com/faceaurus.htm]. [@Holtzman_2011, p. 650]

The link above contains only morphed face images and no further details about the morphing or stimulus preparation procedure.



#### Scriptable Methods

There are several scriptable methods for creating image stimuli, including MatLab, ImageMagick, and GraphicConvertor.

MatLab [@higham2016matlab] is widely used within visual psychophysics. A Google Scholar search for "MatLab face attractiveness" returned 7,440 hits, although the majority of papers I inspected used MatLab to process EEG data, present the experiment, or analyse image color, rather than using MatLab to create the stimuli. "MatLab face perception" generated 80,800 hits, more of which used MatLab to create stimuli.

> The average pixel intensity of each image (ranging from 0 to 255) was set to 128 with a standard deviation of 40 using the SHINE toolbox (function lumMatch) (Willenbockel et al., 2010) in MATLAB (version 8.1.0.604, R2013a). [@visconti2014facilitated, p. 2]


ImageMagick [@imagemagick] is a free, open-source program that creates, edits, and converts images in a scriptable manner. The {magick} R package [@R-magick] allows you to script image manipulations in R using ImageMagick.

> Images were cropped, resized to 150 × 150 pixels, and then grayscaled using ImageMagick (version 6.8.7-7 Q16, x86_64, 2013-11-27) on Mac OS X 10.9.2. [@visconti2014facilitated, p. 2]

GraphicConvertor [@nishimura2000graphicconverter] is typically used to batch process images, such as making images a standard size or adjusting color. While not technically "scriptable", batch processing can be set up in the GUI interface and the saved to a reloadable ".gaction" file. (A search for '"gaction" graphicconvertor' on Google Scholar returned no hits.)

> We used the GraphicConverterTM application to crop the images around the cat face and make them all 1024x1024 pixels. One of the challenges of image matching is to do this process automatically. [@paluszek2019pattern, p.214]

#### Mystery Methods

Many researchers describe image manipulation generically or use "in-house" methods that are not well specified enough for another researcher to have any chance of replicating them.

> Each of the images was rendered in gray-scale and morphed to a common shape using an in-house program based on bi-linear interpolation (see e.g., Gonzalez & Woods, 2002). Key points in the morphing grid were set manually, using a graphics program to align a standard grid to a set of facial points (eye corners, face outline, etc.). Images were then subject to automatic histogram equalization. [@burton2005robust, p. 263]

The reference above [@gonzalez2002digital] has been cited by 2384 papers on Google Scholar, and is a 190-page book. It mentions bilinear interpolation on pages 64--66 in the context of calculating pixel color when resizing images and it's unclear how this could be used to morph shape.


#### Psychomorph/WebMorph

**TO DO**

* templates [@tiddeman2001computer]
* averaging [@benson1993extracting]
* texture [@tiddeman2001prototyping]
* transforming [@rowland1995manipulating]
* webmorph.org [@webmorph]

### Ethical Issues

Research with identifiable faces has a number of ethical issues. This means it is not always possible to share the exact images used in a study. Then it is all the more important for the stimulus construction methods to be clear and reproducible. However, there are other ethical issues outside of image sharing that we feel are important to highlight in an paper discussing the use of face images in research.

The use of face photographs must respect participant consent and personal data privacy. Images that are "freely" available on the internet are a grey area and the ethical issues should be carefully considered by the researchers and relevant ethics board.

Do not use face images in research where there is a possibility of real-world consequences for the pictured individuals. For example, do not post identifiable images of real people on real dating sites without the explicit consent of the pictured individuals for that specific research.

The use of face image analysis should never be used to predict behaviour or as automatic screening. For example, face images cannot be used to predict criminality or decide who should proceed to the interview stage in a job application. This type of application is unethical because the training data is always biased. Face image analysis can be useful for researching what aspects of face images give rise to the *perception* of traits like trustworthiness, but should not be confused with the ability to detect *actual* behaviour. Researchers have a responsibility to consider how their research may be misused in this manner.


### Glossary

```{r glossary}
terms <- tibble::tribble(
  ~Term, ~Definition,
  "template", "a set of landmark points that define shape and the way these are connected with lines; only image with the same template can be averaged or transformed",
  "landmark", "a point that marks corresponding locations on different images",
  "morphing", "blending two or more images to make an image with an average shape andor color",
  "transforming", "changing the shape and/or color of an image by some proportion of a vector that is  defined as the difference between two images",
  "delineation", "the x- and y-coordinates for a specific template that describe an image",
  "lines", "connections between landmarks; these may be used to interpolate new landmarks for morphing",
  "prototype", "an average of faces with similar characteristics, such as expression, gender, age, and/or ethnic group",
  "composite", "an average of more than one face image"
) %>%
  dplyr::arrange(Term)

kableExtra::kable(terms, caption = "Glossary of terms.") %>%
  kable_styling(bootstrap_options = c("striped", "responsive"))
```




## Methods

### Editing

Almost all imagesets start with raw images that need to be cropped, resized, rotated, padded, and/or color normalised. Although many reproducible methods exist to manipulate images in these ways, they are complicated when an image has an associated template, so webmorphR has functions that alter the image and template together (Fig.\ \@ref(fig:editing)).

```{r, echo = TRUE}
orig <- demo_stim("test", 1) # load one demo image
mirrored <- mirror(orig)
cropped <- crop(orig, width = 0.75, height = 0.75)
resized <- resize(orig, 0.75)
rotated <- rotate(orig, degrees = 180)
padded <- pad(orig, 30, fill = "black")
grey <- greyscale(orig)
```

```{r editing, fig.cap="Examples of image manipulations: (A) original image, (B) mirrored, (C) cropped to 75%, (D) resized to 75%, (E) rotated 180 degrees, (F) 30 pixels of black padding added, and (G) greyscale,.", fig.alt="Seven versions of the same average female face with different manipulations applied."}

c(orig, mirrored, cropped, resized, rotated, padded, grey) %>%
  pad(0, 0, 0, 60) %>%
  #draw_tem() %>%
  label(LETTERS, size = 60) %>% 
  plot(nrow = 1)
```


### Delineation

The image manipulations above work best if your raw images start the same size and aspect ratio, with the faces in the same orientation and position on each image. This is frequently not the case with raw images. Image delineation provides a way to set image manipulation parameters relative to face landmarks by marking corresponding points according to a template.

WebMorph's default face template marks 189 points (Fig.\ \@ref(fig:delineate)). Some of these points have very clear anatomical locations, such as point 0 ("left pupil"), while others have only approximate placements and are used mainly for masking or preventing morphing artifacts from affecting the background of images, such as point 147 ("about 2cm to the left of the top of the left ear (creates oval around head)"). Template points numbering is 0-based because PsychoMorph was originally written in Java.

```{r delineate, fig.cap = "Default webmorph FRL template", fig.alt = "Average female face with the shape delineated using 189 white numeric labels and blue lines connecting the numbers."}
demo_stim("composite", "f_multi") %>%
  crop_tem(20) %>%
  resize(2) %>%
  draw_tem(pt.shape = "index", pt.color = "#FFFFFFFF", pt.size = 25) %>%
  plot()
```

The function `tem_def()` retrieves a template definition that includes point names, default coordinates, and the identity of the symmetrically matching point for mirroring or symmetrising images Table\ \@ref(tab:tem-def).

```{r tem-def}
# get all information about a standard template
FRL <- tem_def("FRL")

FRL$points[1:10, 1:5] %>% 
  kableExtra::kable(caption = "The first 10 landmark points of WebMorph's default \"FRL\" template.") %>%
  kable_styling(bootstrap_options = c("striped"))
```


You can automatically delineate faces with a simpler template (Fig.\ \@ref(fig:auto-delin)) using the python package [@face-recognition] or online services provided through the free web platform Face++ [-@faceplusplus].

```{r, echo = TRUE}
# load 5 images with FRL templates
f <- demo_stim("london", "006|038|064|066|135")

# remove templates and auto-delineate with dlib
dlib70_tem <- auto_delin(f[1], "dlib70", replace = TRUE)
dlib7_tem <- auto_delin(f[1], "dlib7", replace = TRUE)

# remove templates and auto-delineate with Face++
fpp106_tem <- auto_delin(f[1], "fpp106", replace = TRUE)
fpp83_tem <- auto_delin(f[1], "fpp83", replace = TRUE)
```

```{r auto-delin, fig.cap="Delineation templates: (A) manual delineation using the FRL template, (B) automatic delineation using the Face++ 106-point template, (C) automatic delineation using the Face++ 83-point template, (D) automatic delineation using the 70-point dlib template, and (E) automatic delineation using the 7-point dlib template.", fig.alt="A West Asian female face showing dots and lines marking the full face (top row), three reduced template marking only the jaw from ear to ear and internal facial features, and one marking only the eyes and nose."}

c(f[1], fpp106_tem, fpp83_tem, dlib70_tem, dlib7_tem) %>%
  draw_tem() %>%
  label(LETTERS) %>%
  plot()
```

A study comparing the accuracy of four common measures of face shape (sexual dimorphism, distinctiveness, bilateral asymmetry, and facial width to height ratio) between automatic and manual delineation concluded that automatic delineation had higher replicability and good correlations with manual delineation [@jones2021facial]. However, around 2% of images had noticeably inaccurate automatic delineation, which should be screened for by outlier detection and visual inspection. 

While automatic delineation has the advantage of being very fast and generally more replicable than manual delineation, it is more limited in the areas that can be described. Typically, automatic face detection algorithms outline the lower face shape and internal features of the face, but don't define the hairline, hair, neck, or ears. Manual delineation of these can greatly improve stimuli created through morphing or transforming (Fig.\ \@ref(fig:avg-comp)).

```{r avg-comp, fig.cap="Averages of 5 images made using (A) the full 189-point manual template and (B) the reduced 106-point automatic template.", fig.alt="Averages of the 5 West Asian female faces. The right average has blurrier features then the left one, especially around the ears, neck and hairline."}
frl_avg <- avg(f)
fpp_avg <- avg(fpp_tem)

c(frl_avg, fpp_avg) %>% 
  label(LETTERS) %>%
  plot()
```

### Facial Metrics

Once you have images delineated, you can use the x- and y-coordinates to calculate various facial-metric measurements (Table\ \@ref(tab:metrics)). You can calculate the distance between any two points, such as the pupil centres, or use a more complicated formula, such as the face width-to-height ratio from Lefevre et al. [-@lefevre2013telling]. 

```{r, echo = TRUE}
# inter-pupillary distance between points 0 and 1
ipd <- metrics(f, c(0, 1))

# face width-to-height ratio
# width = maximum bizygomatic width
# height = top of upper lip to highest eyelid point
fwh <- metrics(f, "abs(max(x[113],x[112],x[114])-min(x[110],x[111],x[109]))/abs(y[90]-min(y[20],y[25]))")
```

```{r metrics}
data.frame(
  face = names(f),
  x0 = metrics(f, "x[0]"),
  y0 = metrics(f, "y[0]"),
  x1 = metrics(f, "x[1]"),
  y1 = metrics(f, "y[1]"),
  ipd = ipd,
  fwh = fwh
) %>%
  kable(caption = "Facial metric measurements.", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "responsive"))
```

While it is *possible* to calculate metrics such as width-to-height ratio from 2D face images, this does not mean it is a good idea. Even on highly standardized images, head tilt can have large effects on such measurements. When image qualities such as camera type and head-to-camera distance are not standardized, facial metrics are meaningless at best [@tvrebicky2016focal].

### Alignment

If your image set isn't highly standardised, you probably want to crop, resize and rotate your images to get them all in approximately the same orientation on images of the same size. There are several reproducible options, each with pros and cons. 

One-point alignment (Fig.\ \@ref(fig:norm-comp)A) doesn't rotate or resize the image at all, but aligns one of the delineation points across images. This is ideal when you know that your camera-to-head distance and orientation was standard (or meaningfully different) across images and you want to preserve this in the stimuli, but you still need to get them all in the same position and image size.

Two-point alignment (Fig.\ \@ref(fig:norm-comp)B) resizes and rotates the images so that two points (usually the centres of the eyes) are in the same position on each image. This will alter relative head size such that people with very close-set eyes will appear to have larger heads than people with very wide-set eyes. This technique is good for getting images into the same orientation when you didn't have any control over image rotation and camera-to-head distance of the original photos. 

Procrustes alignment (Fig.\ \@ref(fig:norm-comp)C) resizes and rotates the images so that each delineation point is as aligned as possible across all images. This can obscure meaningful differences in relative face size (e.g., a baby's face will be as large as an adult's), but can be superior to two-point alignment. While this requires that the whole face be delineated, you can use a minimal template such as a face outline or the Face++ auto-delineation to achieve good results. While not available in webmorph.org, procrustes alignment is available in webmorphR. 

You can very quickly delineate an image set with a custom template using the `quick_delin()` function in webmorphR if auto-delineation doesn't provide suitable points.

```{r, echo = TRUE}
# one-point alignment
onept <- align(f, pt1 = 55, pt2 = 55,
               x1 = width(f)/2, y1 = height(f)/2,
               fill = "dodgerblue")

# two-point alignment
twopt <- align(f, pt1 = 0, pt2 = 1, fill = "dodgerblue")

# procrustes alignment
proc <- align(f, pt1 = 0, pt2 = 1, procrustes = TRUE, fill = "dodgerblue")
```

```{r norm-comp, fig.cap="Original images with different alignments. (A) One-point alignment placing the bottom of the nose point in the centre of the image. (B) Two-point alignment placing the eye centre points in the same position as the average image. (C) Procrustes alignment moved, rotated, and resized all images to most closely match the average face. A blue background was used to highlight the difference here, but normally a colour matching the image background would be used or the images would be cropped.", fig.alt="Five West Asian female faces on each row, with the alignments described in the caption."}

plot_rows(A = onept, B = twopt, C = proc)
```



### Masking

Oftentimes, researchers will want to remove the background, hair, and clothing from an image to avoid confounds. For example, the presence versus absence of hairstyle information can reverse preferences for masculine versus feminine male averages [@debruine2006correlated].

The "standard oval mask" has enjoyed widespread popularity because it is straightforward to add to images using programs like PhotoShop. WebmorphR's `mask_oval()` function allows you to set oval boundaries manually (Fig.\ \@ref(fig:mask)A) or in relation to minimum and maximum template coordinates for each face (Fig.\ \@ref(fig:mask)B). An arguably better way to mask out hair, clothing and background from images is to crop around the curves defined by the template (Fig.\ \@ref(fig:mask)C). 

```{r, echo = TRUE}
# standard oval mask
bounds <- list(t = 200, r = 400, b = 300, l = 400)
oval <- mask_oval(f, bounds, fill = "dodgerblue")

# template-aware oval mask
oval_tem <- f %>%
  subset_tem(features("gmm")) %>% # remove external points
  mask_oval(fill = "dodgerblue") # oval boundaries to max and min template points

# template aware mask
masked <- mask(f, c("face", "neck", "ears"), fill = "dodgerblue")
```

```{r mask, fig.cap = "Images masked with (A) an oval defined by image coordinates, (B) an oval defined by the minimum and maximum x- and y-coordinates of template points, or (C) to include face, ears and neck.", fig.alt="Five West Asian female faces on each row, with the masking described in the caption."}
plot_rows(A = oval, B = oval_tem, C = masked)
```

### Averaging

Creating average images (also called composite or prototype images) through morphing can be a way to visualise the differences between groups [@burton2005robust], manipulate averageness [@Little_2011], or create prototypical faces for image transformations. 

Averaging faces with texture [@tiddeman2001prototyping;@tiddeman2005towards] makes composite image look more realistic (Fig.\ \@ref(fig:avg-texture)A). However, averages created without texture averaging look smoother and may be more appropriate for transforming color (Fig.\ \@ref(fig:avg-texture)B). 

```{r, echo = TRUE}
avg_tex <- avg(f, texture = TRUE)
avg_notex <- avg(f, texture = FALSE)
```


```{r avg-texture, fig.cap="An average of 5 faces created (A) with texture averaging and (B) without.", fig.alt="The average female West Asian face. The Top row shows the full face and the bottom row is a close-up of the eyes. The left image has cripsly defined features and skin texture, such as wrinkles and visible eyebrow hairs, while the right image has blurrier features, smoother skin, and individual hairs are not visible."}
# close up on the eyes
eyes <- c(avg_tex, avg_notex) %>% 
  align(x1=100, y1 = 100, x2 = 300, y2 = 100, width = 400, height = 150)
  
c(avg_tex, avg_notex, eyes) %>% 
  label(c("A", "B", "", "")) %>%
  resize(width = width(avg_tex)) %>%
  plot(nrow = 2)
```

### Transforming

Transforming alters the appearance of one face by some proportion of the differences between two other faces. This technique is distinct from morphing. For example, you can transform a face in the dimension of sexual dimorphism by calculating  the shape and color differences between a prototype female face (Fig.\ \@ref(fig:trans-vs-morph)A) and a prototype male face (Fig.\ \@ref(fig:trans-vs-morph)B). If you morph an individual female face with these images, you get faces that are halfway between the indivual and prototype faces (Fig.\ \@ref(fig:trans-vs-morph)C,D). If you transform the individual face by 50% of the prototype differences, you get feminised and masculinized versions of the individual face (Fig.\ \@ref(fig:trans-vs-morph)E,F).

```{r}
# load individual and composite images
individual <- demo_stim("london", "019")
averages <- demo_stim("composite", "white")

# mask the composites so they only transform colour in the face
masked <- mask(averages, "face")

# make masc and fem versions of the individual
sexdim <- trans(trans_img = individual,
              from_img = masked$f_white,
              to_img  = masked$m_white,
              shape = c(fem = -0.5, masc = 0.5), 
              color = c(fem = -0.5, masc = 0.5))

# average the individual face with male and female averages
masc_morph <- c(individual, averages$m_white) %>% avg()
fem_morph <- c(individual, averages$f_white) %>% avg()
```

```{r trans-vs-morph, fig.cap="Morphing versus transforming: (A, B) composite images, (C, D) averages of the composites with an individual image, (E, F) transforms of an individual image along the male-female continuum.", fig.alt="Six faces in three columns/two rows. The first column is average White faces, female on top and male on the bottom. The second column is the average of the faces in the first column and an individual White female face; the top image looks like a more average version of the individual face and the bottom image looks like an androgynous face. The third column is transforms of the individual female face; the top image looks more like a more feminine version of the individual face and the bottom image looks like a male version of the individual face."}

c(averages, fem_morph, masc_morph, sexdim) %>%
  label(LETTERS) %>%
  plot(nrow = 2,byrow = FALSE)
```

If, for example, the individual female face was more feminine than the average female face, morphing with the average female face produces an image that is *less* feminine than the original individual, while transforming along the male-female dimension produces and image that is always *more* feminine than the original. Morphing with a prototype also results in an image with increased averageness, while transforming maintains individually distinctive features.


Transforming also allows you to manipulate shape and colour independently (Fig.\ \@ref(fig:trans-shape-color)).

```{r}
masc <- trans(trans_img = individual,
              from_img = masked$f_white,
              to_img  = masked$m_white,
              shape = c(shape = 0.5, color = 0, both = 0.5), 
              color = c(shape = 0, color = 0.5, both = 0.5))
```

```{r trans-shape-color, fig.cap="Transforming shape and color independently: (A) original individual image, (B) shape only, (C), color only, (D) both shape and color."}
c(individual, masc) %>%
  label(LETTERS) %>%
  plot(nrow = 2)
```


### Symmetrising

Although a common technique [e.g., @Mealey_1999], left-left and right-right mirroring (Fig.\ \@ref(fig:mirror-sym)) is not recommended for investigating perceptions of facial symmetry. This is because this method typically produces unnatural images for any face that isn't already perfectly symmetric. For example, if the nose does not lie in a perfectly straight line from the centre point between the eyes to the centre of the mouth, then one of the mirrored halves will have a much wider nose than the original face, while the the other half will have a much narrower nose than the original face. In extreme cases, one mirrored version can end up with three nostrils and the other with a single nostril. 

```{r mirror-sym, fig.cap = "Left-left (top) and right-right (bottom) mirrored images. The code for making these images is in the supplemental materials, but we only recommend using this method to demonstrate how misleading it is.", fig.alt="The five West Asian women "}
# make eye points exactly horizontal
hzeyes <- horiz_eyes(f)

# calculate midpoint of eyes for each image
pts <- get_point(hzeyes, pt = 0:1)
midpoint <- mean(pts$x)

# crop and mirror images
left_side <- crop(hzeyes, width = midpoint, x_off = 0)
left_mirror <- mirror(left_side)
right_side <- crop(hzeyes, width = width(hzeyes)-midpoint, x_off = midpoint)
right_mirror <- mirror(right_side)

# paste images together
left_left <- mapply(function(ls, rs) {
  c(ls, rs) %>% plot(padding = 0)
}, left_side, left_mirror) %>%
  crop(width(., "max"), height(., "max"))


right_right <- mapply(function(ls, rs) {
  c(ls, rs) %>% plot(padding = 0)
}, right_mirror, right_side) %>%
  crop(width(., "max"), height(., "max"))

plot_rows(left_left, right_right)
```

A morph-based technique is a more realistic way to manipulate symmetry [@Little_2011,@little2001self,@paukner2017capuchin]. It preserves the individual's characteristic feature shapes and avoids the problem of having to choose an axis of symmetry on a face that isn't perfectly symmetrical. In this method, the original face is mirror-reversed and each template point is re-labelled. The original and mirrored images are averaged together to create a perfectly symmetric version of the image that has the same feature widths as the original face (Fig.\ \@ref(fig:morph-sym)). You can also use this symmetric version to create asymmetric versions of the original face through transforming, exaggerating the differences between the original and the symmetric version.

```{r, echo = TRUE}
sym_both <- symmetrize(f)
sym_shape <- symmetrize(f, color = 0)
sym_color <- symmetrize(f, shape = 0)
sym_anti <- symmetrize(f, shape = -1.0, color = 0)
```

```{r morph-sym, fig.cap = "Images with different types of symmetry: (A) symmetric shape and color, (B) symmetric shape, (C) symmetric color, (D) asymmetric shape."}
plot_rows(A = sym_both, B = sym_color, C = sym_shape, D = sym_anti)
```


## Case Study

### London Face Set

We will use the open-source, CC-BY licensed image set, the Face Research Lab London Set [@FRL_London]. Images are of 102 adults whose pictures were taken in London, UK, in April 2012 for a project with Nikon camera (Fig.\ \@ref(fig:london-set)). All individuals were paid and gave signed consent for their images to be "used in lab-based and web-based studies in their original or altered forms and to illustrate research (e.g., in scientific journals, news media or presentations)." 

```{r london-set, fig.cap = "The 102 neutral front faces in the London Face Set."}

london <- demo_stim("london") %>%
  resize(0.5) %>% # makes the demo run faster, change for final
  add_info(stimsets::london_info)

london %>% plot(nrow=6, maxwidth = 2000)
```

```{r}
smiling <- demo_stim("smiling") %>%
  resize(0.5) %>% # makes the demo run faster, change for final
  add_info(stimsets::london_info)
```

Each subject has one smiling and one neutral pose. For each pose, 5 full colour images were simultaneously taken from different angles: left profile, left three-quarter, front, right three-quarter, and right profile, but we will only use the front-facing images in the examples below. These images were cropped to 1350x1350 pixels and the faces were manually centered (Lisa did this many years ago before we made the tools in this paper). The neutral front images have template files that mark out 189 coordinates delineating face shape for use with Psychomorph or WebMorph.

### Protoypes

The first step for many types of stimuli is to create prototype faces for some categories, such as expression or gender. The faces that make up these averages should be matched for other characteristics that you want to avoid confounding with the categories of interest, such as age or ethnicity. Here, we will choose 5 Black female faces, automatically delineate them, align the images, and create neutral and smiling prototypes (Fig.\ \@ref(fig:emo-avg)).

```{r, echo = TRUE}
# select the relevant images and auto-delineate them
neu_orig <- subset(london, face_gender == "female") %>% 
  subset(face_eth == "black") %>% subset(1:5) %>%
  auto_delin(replace = TRUE)

smi_orig <- subset(smiling, face_gender == "female") %>% 
  subset(face_eth == "black") %>% subset(1:5) %>%
  auto_delin(replace = TRUE)

# align the images
aligned <- c(neu_orig, smi_orig) %>%
  align(procrustes = TRUE, patch = TRUE) %>%
  crop(.6, .8, y_off = 0.05)

neu <- subset(aligned, 1:5)
smi <- subset(aligned, 6:10)

neu_avg <- avg(neu, texture = FALSE)
smi_avg <- avg(smi, texture = FALSE)
```

```{r emo-avg, fig.cap="Average and individual neutral and smiling faces."}
c(neu_avg,
  plot(neu, external_pad = 0, ncol = 1), 
  plot(smi, external_pad = 0, ncol = 1),
  smi_avg
) %>% 
  resize(height = height(neu_avg)) %>%
  plot()
```

### Emotion Continuum

Once you have two prototype images, you can set up a continuum that morphs between the images and even exaggerates beyond them (Fig.\ \@ref(fig:continuum)). Note that some exaggerations beyond the prototypes can produce impossible shape configurations, such as the negative smile, where the open lips from a smile go to closed at 0% and pass through each other at negative values.

```{r, echo = TRUE}
steps <- continuum(neu_avg, smi_avg, from = -0.5, to = 1.5, by = 0.25)
```

```{r continuum, fig.cap="Continuum from -50% to +150% smiling."}
lab <- paste0(seq(-0.5, 1.5, .25) * 100, "%")
steps %>%
  label(lab) %>%
  plot(nrow = 1)
```


### Sexual dimorphism transform

We can use the full templates to create sexual dimorphism transforms from neutral faces. Repeat the process above for 5 male and 5 female neutral faces, skipping the auto-delineation (Fig.\ \@ref(fig:sexdim-avg)).

```{r, echo = TRUE}
# select the relevant images
f_orig <- subset(london, face_gender == "female") %>% 
  subset(face_eth == "black") %>% subset(1:5)

m_orig <- subset(london, face_gender == "male") %>% 
  subset(face_eth == "black") %>% subset(1:5)

# align the images
aligned <- c(f_orig, m_orig) %>%
  align(procrustes = TRUE, patch = TRUE) %>%
  crop(.6, .8, y_off = 0.05)

f <- subset(aligned, 1:5)
m <- subset(aligned, 6:10)

f_avg <- avg(f, texture = FALSE)
m_avg <- avg(m, texture = FALSE)
```

```{r sexdim-avg, fig.cap="Average and individual female and male faces."}
c(f_avg,
  plot(f, external_pad = 0, ncol = 1), 
  plot(m, external_pad = 0, ncol = 1),
  m_avg
) %>% 
  resize(height = height(f_avg)) %>%
  plot()
```

Next, transform each individual image using the average female and male faces as transform endpoints (Fig.\ \@ref(fig:sexdim-transform)).

```{r, echo = TRUE}
# use a named vector for shape to automatically rename the images
sexdim <- trans(
  trans_img = c(f, m),
  from_img = f_avg,
  to_img = m_avg,
  shape = c(fem = -.5, masc = .5)
)
```


```{r sexdim-transform, fig.cap="Versions of individual faces with (A) 50% feminised shape and (B) 50% masculinized shape."}

plot_rows(A = subset(sexdim, "_fem"),
          B = subset(sexdim, "_masc"))

```


### Self-resemblance transform

Much of my own research involves creating "virtual siblings" for participants to test how they perceive and behave towards strangers with phenotypic kinship cues [@DeBruine_2004PRSLB;@DeBruine_2005PRSLB;@DeBruine_2011PNAS]. As discussed in detail in DeBruine et al. [-@DeBruine_2008ASB], while morphing techniques are sufficient to create same-gender virtual siblings, transforming techniques are required to make other-gender virtual siblings without confounding self-resemblance with androgyny (Fig.\ \@ref(fig:virtual-sibs)).

```{r, echo = TRUE}
virtual_sis <- trans(
  trans_img = f_avg,   # transform an average female face
  shape = 0.5,         # by 50% of the shape differences
  from_img = m_avg,    # between an average male face
  to_img = m) %>%      # and individual male faces
  mask(c("face", "neck","ears")) 

virtual_bro <- trans(
  trans_img = m_avg,   # transform an average male face
  shape = 0.5,         # by 50% of the shape differences
  from_img = m_avg,    # between an average male face
  to_img = m) %>%      # and individual male faces
  mask(c("face", "neck","ears"))
```

```{r virtual-sibs, fig.cap="Creating virtual siblings: (A) original images, (B) virtual brothers, (C) virtual sisters."}
plot_rows(A = crop_tem(m), 
          B = crop_tem(virtual_bro), 
          C = crop_tem(virtual_sis))
```

### Labels

Many social perception studies require labelled images, such a minimal group designs. You can add custom labels and superimpose images on stimuli (Fig.\ \@ref(fig:label-comp)).

```{r}
# download, resize and save flag images

dir.create("images/flags", FALSE)

magick::image_read_svg("https://upload.wikimedia.org/wikipedia/commons/1/10/Flag_of_Scotland.svg") %>%
  magick::image_resize(magick::geometry_size_pixels(75)) %>%
  magick::image_write("images/flags/saltire.png")

magick::image_read_svg("https://upload.wikimedia.org/wikipedia/commons/a/a9/Flag_of_Wales_%281959%E2%80%93present%29.svg") %>%
  magick::image_resize(magick::geometry_size_pixels(75)) %>%
  magick::image_write("images/flags/ddraig.png")
```


```{r, echo = TRUE}

flags <- read_stim("images/flags")

ingroup <- f %>%
  # pad 10% at the top with matching color
  pad(0.1, 0, 0, 0, patch = TRUE) %>% 
  label("Scottish", "north", "+0+10") %>%
  image_func("composite", flags$saltire$img, 
              gravity = "northeast", offset = "+10+10")

outgroup <- f %>%
  pad(0.1, 0, 0, 0, patch = TRUE) %>% 
  label("Welsh", "north", "+0+10") %>%
  image_func("composite", flags$ddraig$img, 
             gravity = "northeast", offset = "+10+10")
```

```{r label-comp, fig.cap="Stimuli with text labels and superimposed images."}
plot_rows(ingroup, outgroup)
```




## Discussion

* head position in 2D images
  * morphometics
  * facefuns
* Natural vs standardised source images
  * right image for the question
* Averaging is N=1 https://debruine.github.io/posts/composite-images/



\newpage

## References

We used `r cite_r("r-references.bib")` to produce this manuscript.

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
